---
title: "Introduction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction


In a span of a few months, the COVID-19 pandemic caused by the coronavirus radically transformed the lives of many and ultimately changed the way various industries, businesses, and/or sectors have functioned. Among those affected by the pandemic have been colleges universities, specifically college advancement and fundraising.  As the pandemic swept the world, the higher education institutions were forced to deal with numerous issues surrounding income losses (How COVID-19 Affected Higher Education, 2021).  One issue was a loss of income due to virtual learning. With education and classes being remote, most on-campus housing arrangements were vacated. And due to a lack of students on campus, any activity that would’ve otherwise gained income from students being on campus was now rendered obsolete.  To combat this issue, colleges have to adopt new methods of procuring funds.  One method that has been cited as being effective is through adopting new donor engagement strategies (How COVID-19 Affected Higher Education, 2021). Understanding which factors have driven donations in the past could help universities better target likely donors can help improve fundraising and help mitigate or recover income loss moving forward. 

With that in mind, this project aimed to determine what those factors could be.  To do this, a data set called “University Advancement, Donations, and Giving” was obtained from Tableau. The data set has 3914 observations across nine variables. This data set is limited as it does not explore various macro and micro-level variables such as environmental(political-sociological) influences, endowment, population, private or public standing, type of school (undergraduate, graduate, professional) that could have helped inform our analysis. Nonetheless, insights were able to be drawn from the nine variables provided in the data set. 


## Description of Data

# Source Data

Using the data, this project sought to determine what factors affected the amount of money donated to universities from 2010 - 2015. The donation amount will be compared across the nine variables and tested to see if a difference exists. As mentioned before, the dataset was obtained from Tableau's sample data and required minimal preprocessing. 

```{r echo=FALSE}
library(readr)
UniversityDonations <- read.csv("/Users/chim/Downloads/UniversityDonations.csv")
str(UniversityDonations)
```
Allocation sub_category: Donor Type
AGR: College of Agriculture & Natural Science
ALM: Alumni Association
ANN: University Annual Fund
ART: College of Arts & Sciences
ATH: Athletics
BIZ: College of Business
COM: College of Communication Arts & Sciences
DIV: Diversity Fund
EDU: College of Education
ENG: College of Engineering
FCE: Faculty Chair Endowments
HON: Honors Program
ISF: In-state Scholarship Fund
LIB: University Library
MSF: Minority Scholarship Fund
MSM: Campus Museums
MUS: College of Music
NAT: College of Natural Science
NUR: College of Nursing
PAR: Parents Association
POL: College of Political Science
RCR: Campus Recreations
TST: Trustees Fund
VET: College of Veterinary Medicine

City and State: Location of the university the money was allocated to

Gift_Allocation: Form of donation/funds

Major: Department 

College: Discipline 

Amount: amount of money allocated

Year: year gift was allocated

Prospect.ID - Id of recipient 


Before our exploratory data analysis, it was necessary to change all variables but donation amount to factors as they were categorical variables and it was necessary to remove any outliars. In addition, the contents of the 'Allocation sub_category' column were abbreviated as the contents were too long.  For our exploratory data analysis, we ignored the 'Prospect.ID' column as they are variables with no relation to the donation amount. And we ran the donation amount through a logarithmic function to allow for better visualizations and plots.
